---
title: "MUSSER-GRACE-HOMEWORK-01"
author: "Grace Musser"
date: "2/5/2019"
output: html_document
---

Module 7 Challenges 

CHALLENGE:
First, create a vector of 1 word character strings comprising the first line of the Gettysburg address: “Four score and seven years ago our fathers brought forth on this continent, a new nation, conceived in Liberty, and dedicated to the proposition that all men are created equal.”
Then, extract every third element of this vector.

HINT 1: Take a look at the {stringr} package and the function str_split() to see if you can easily divide a single string into a vector of substrings.
HINT 2: If you’re ambitious, the following function, gsub("[[:punct:]]","",*string*), applied to *string*, will remove punction and special characters from that string.

```{r}
Lincoln<-c("Four score and seven years ago our fathers brought forth on this continent, a new nation, conceived in Liberty, and dedicated to the proposition that all men are created equal.")

Lincoln_split<-c(str_split(Lincoln, " ", simplify = TRUE))
Lincoln2<-Lincoln_split[seq(from =1, to=30, by =3)]
print(Lincoln2)
no_punct<-gsub(("[[:punct:]]"), "", Lincoln)
print(no_punct)
```

CHALLENGE:
Given the matrix, m, above, extract the 2nd, 3rd, and 6th columns and assign them to the variable x
Given the matrix, m, above, extract the 6th to 8th row and assign them to the variable x
Given the matrix, m, above, extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x

```{r}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
print(m)
x <- m[6:8, ]
print(x)
x<-m[2:6,2:9]
print(x)
```

CHALLENGE:
Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1 (i.e., a descending series)
Given this matrix, what would the following return?

a[1, 1, 1, 2]
a[2, 3, 2, ]
a[1:5, 1:5, 3, 3]

```{r}
a<-array(data = 400:1, dim=c(5,5,4,4))
print(a)
```

```{r}
a[1, 1, 1, 2]
```

```{r}
a[2, 3, 2, ]
```

```{r}
a[1:5, 1:5, 3, 3]
```


CHALLENGE:
Create a list representing the (simplified) primate taxonomy outlined below. HINT: you can use lists as elements in a list.


```{r}

```



```{r}
Suborders <- list('Haplorhini' , 'Strepsirhini')
Lsuborders <- list('Anthropoidea', 'Tarsioidea', 'Lorisoidea', 'Lemuroidea')
LLsuborders <- list('Platyrrhihni', 'Catarrhihni')
Plat_fam<-list('Cebidae', 'Atelidae', 'Pitheciidae')
Cat_fam<-list('Cercopithecidae', 'Hylobatidae', 'Hominidae')
Tars_fam<-list('Tarsiidae')
Loris_fam<-list('Lorisidae', 'Galagidae')
Lemur_fam<-list('Cheirogaleidae', 'Lepilemuridae', 'Indriidae', 'Lemuridae', 'Daubentoniidae')

Platyrrhini<-list(LLsuborders[[1]], list(Plat_fam))
Catarrhini<-list(LLsuborders[[2]], list(Cat_fam))
Anthropoidea<-list(Lsuborders[[1]], Catarrhini, Platyrrhini)
Tarsoidea<-list(Lsuborders[[2]], Tars_fam)
Haplorhini<-list(Anthropoidea, Tarsoidea)


Lorisoidea<-list(Lsuborders[[3]], list(Loris_fam))
Lemuroidea<-list(Lsuborders[[4]], list(Lemur_fam))
Strepsirhini<-list(Lorisoidea, Lemuroidea)

Primates<-list(Haplorhini, Strepsirhini)
Primates

```


CHALLENGE:
Store the following numbers as a 5 x 3 matrix: 3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2. Be sure to fill the matrix ROWWISE.

```{r}
m <- matrix(data = c(3, 0, 1, 23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), nrow = 5, ncol = 3, byrow = TRUE)
m
```


Then, do the following:
Coerce the matrix to a data frame.

```{r}
df <- data.frame(m)
df
```

As a data frame, coerce the second column to be logical-valued (Boolean)
```{r}
df$X2 <- as.logical(df$X2)
df
class(df$X2)
```


As a data frame, coerce the third column to be factor-valued

```{r}
df$X3 <- as.factor(df$X3)
df
class(df$X3)
```

When you are done, use the str() command to show the data type for each variable in your dataframe.

```{r}
str(df)
```



