---
title: "Musser_Grace-ADA-HOMEWORK-02"
author: "Grace Musser"
date: "2/26/2019"
output: html_document
---
Poisson--use for independently occuring, discrete random variables over intervals. eg DNA mutation accumulation model

Question 1

```{r}
interval<-2
session_hour_total<-14
l<-15 #set lambda to 15 as it is the mean number of occurances per interval
```

1a. 
```{r}
x<-8 #x is 8 calls
prob1<-ppois(x,lambda=l,lower.tail=FALSE) #P[X>x]
prob1
```

1b.
```{r}
x<-0 #x is no calls
prob2<-dpois(x,l) #P[X=x]
prob2
```


1c.
```{r}
x<-3 #x is 3 calls
prob3<-dpois(x,l) #P[X=x]
prob3
```

1d.
```{r}
x<-0:30 #set range of x
prob4<-dpois(x=x, lambda=(15)) #use the poisson dist., lambda is 15
barplot(prob4, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)", main = paste0("Probability Mass Function\nlambda = ", 
    l)) #all of this code plots the histogram
```

1e/1f. It is different from the one plotted above as the distribution is not normal and the results are not as predictable.
```{r}
seed<-rpois(104, 15) #simulate poisson distribution
manipulate(hist(seed, main = paste("Histogram of Simulated Calls", sep = ""), xlab = "calls", xlim=c(0,30), ylab = "time", n = slider(0, 
    30, initial = 1, step = 1 ))) #all of this codes relates to histogram that can be manipulated using manipulate package
```






Question 2

2a.
```{r}
library(curl) #read in data frame after loading curl package
f <- f <- curl("https://raw.githubusercontent.com/difiore/ADA-2019/master/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```
```{r}
n<-1000 #n is a thousand people

#function to calculate pop st dev
sd.p=function(x){
  sd(x)*sqrt((length(x)-1)/length(x))
  }

mean_height<- mean(d$height) #get population means of all variables
x<-d$height
s<-sd.p(x) #get standard dev of population for all variables
mean_height
s

mean_weight<- mean(d$weight)
x<-d$weight
s<-sd.p(x)
mean_weight
s

mean_age<- mean(d$age)
x<-d$age
s<-sd.p(x)
mean_age
s

mean_z<- mean(d$zombies_killed)
x<-d$zombies_killed
s<-sd.p(x)
mean_z
s

mean_edu<- mean(d$years_of_education)
x<-d$years_of_education
s<-sd.p(x)
mean_edu
s

```

2b.
```{r}
library(ggplot2) #load ggplot

```
```{r}
#plot gender by height
height_boxplot <- ggplot(d, aes(x=d$gender, y=d$height)) + 
  geom_boxplot()
height_boxplot
```

```{r}
#plot gender by weight
weight_boxplot <- ggplot(d, aes(x=d$gender, y=d$weight)) + 
  geom_boxplot()
weight_boxplot
```

```{r}
#plot gender by age
age_boxplot <- ggplot(d, aes(x=d$gender, y=d$age)) + 
  geom_boxplot()
age_boxplot
```

```{r}
#plot gender by zombies killed
z_boxplot <- ggplot(d, aes(x=d$gender, y=d$zombies_killed)) + 
  geom_boxplot()
z_boxplot
```

```{r}
#plot gender by years of education
edu_boxplot <- ggplot(d, aes(x=d$gender, y=d$years_of_education)) + 
  geom_boxplot()
edu_boxplot
```

2c.
```{r}
par(mfrow=c(1,2)) #change available plot spcae
ggplot(d, aes(x=d$age, y=c(d$weight))) + geom_point() #create a scatterplot
ggplot(d, aes(x=d$age, y=c(d$height))) + geom_point()
```

2d. Age and height appear normally distributed, whereas years of education and the number of zombies killed are not and appear to be drawn from a Poisson distribution.
```{r}
hist(d$age) #histogram
qqnorm(d$age) #qqplot
hist(d$height)
qqnorm(d$height)
hist(d$zombies_killed)
qqnorm(d$zombies_killed)
hist(d$years_of_education)
qqnorm(d$years_of_education)

```

2e. 
```{r}

#sample from the old dataframe (1 iteration, 30 total) to create new dataframe, do not use replacement
mysample <-d[sample(1:nrow(d), 30,
   replace=FALSE),]

mysample

#calculate sample mean, sample st dev, sample error and confidence intervals for height:
sample_mean<-mean(mysample$height)
sample_mean
sample_SD<-sd(mysample$height)
sample_SD
sem<-sample_SD/sqrt(length(mysample$height))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence

#do the same for the other variables...
sample_mean<-mean(mysample$weight)
sample_mean
sample_SD<-sd(mysample$weight)
sample_SD
sem<-sample_SD/sqrt(length(mysample$weight))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence

sample_mean<-mean(mysample$age)
sample_mean
sample_SD<-sd(mysample$age)
sample_SD
sem<-sample_SD/sqrt(length(mysample$age))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence

sample_mean<-mean(mysample$zombies_killed)
sample_mean
sample_SD<-sd(mysample$zombies_killed)
sample_SD
sem<-sample_SD/sqrt(length(mysample$zombies_killed))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence

sample_mean<-mean(mysample$years_of_education)
sample_mean
sample_SD<-sd(mysample$years_of_education)
sample_SD
sem<-sample_SD/sqrt(length(mysample$years_of_education))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence


```




```{r}
#check histograms of each variable sampled
hist(mysample$height)
hist(mysample$weight)
hist(mysample$age)
hist(mysample$zombies_killed)
hist(mysample$years_of_education)
```

2f. Height, weight and age appear normally distributed but zombies killed and years of education do not. The latter two appear to bedrawn from a Poisson distribution. The standard error appears smaller in 2f in general.
```{r}
#do the same thing with 99 iterations...
mysample2 <-d[sample(99:nrow(d), 30,
   replace=FALSE),]
mysample2

sample_mean<-mean(mysample2$height)
sample_mean
sample_SD<-sd(mysample2$height)
sample_SD
sem<-sample_SD/sqrt(length(mysample2$height))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence

sample_mean<-mean(mysample2$weight)
sample_mean
sample_SD<-sd(mysample2$weight)
sample_SD
sem<-sample_SD/sqrt(length(mysample$weight))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence

sample_mean<-mean(mysample2$age)
sample_mean
sample_SD<-sd(mysample2$age)
sample_SD
sem<-sample_SD/sqrt(length(mysample$age))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence

sample_mean<-mean(mysample2$zombies_killed)
sample_mean
sample_SD<-sd(mysample2$zombies_killed)
sample_SD
sem<-sample_SD/sqrt(length(mysample$zombies_killed))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence

sample_mean<-mean(mysample2$years_of_education)
sample_mean
sample_SD<-sd(mysample2$years_of_education)
sample_SD
sem<-sample_SD/sqrt(length(mysample$years_of_education))
sem
confidence<-c(mean(x)-2*sem,mean(x)+2*sem)
confidence
```

```{r}
hist(mysample2$height)
hist(mysample2$weight)
hist(mysample2$age)
hist(mysample2$zombies_killed)
hist(mysample2$years_of_education)
```

